---
output:
  github_document:
    toc: FALSE
---

```{r, echo = FALSE}
# options(knitr.kable.NA = '',
#         digits = 1,
#         width = 80)
set.seed(496)
library(WBdif)
```


# WBdif

**WBdif** contains functions to detect differential item functioning (DIF), estimate treatment effects, evaluate the robustness of treatment effects in the presence of DIF, and	produce reports summarizing the results. Development of **WBdif** was motivated by questions regarding the psychometric properties of the early childhood development assessments used by World Bank to evaluate intervention impacts.

The 3 main functions in **WBdif** are [`dif_analysis()`](#analysis), [`robustness()`](#robustness), and [`dif_report()`](#report).



## Installation


```{r, warning=FALSE, message=FALSE, eval=FALSE}
install.packages("remotes")
remotes::install_github("knickodem/WBdif")
```


## Features

Say you have an assessment with both dichotomous and polytomous items and you want to know:

  1. Do any items show DIF by gender?
  2. What method should I use to detect DIF?
  3. What is the treatment effect of the intervention?
  4. Does the treatment effect change when accounting for DIF-y items?
  
The `dif_analysis()` function conducts a DIF analysis wiht any or all of the methods: Visual inspection via LOESS curves, Mantel-Haenszel, logistic regression, and item response theory. `dif_report()` then displays the results allowing you to compare the findings from each method. 

### Preparing the Data

Prior to using any of the 3 main functions, an optional first step is to use the `dif_prep()` function to prepare the data.  

### Differential Item Functioning Analysis {#analysis}


### Treatment Effect Robustness {#robustness}

### Reporting {#report}

## Support and Suggestions

If you encounter any issues or have a suggestion for additional features, please file a [Github issue](https://github.com/knickodem/WBdif/issues) or contact us via email.