---
output:
  github_document:
    toc: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# WBdif

**WBdif** contains functions to detect differential item functioning (DIF), estimate unconditional and conditional treatment effects, evaluate the robustness of treatment effects in the presence of DIF, and	produce reports summarizing the results. Development of **WBdif** was motivated by questions regarding the psychometric properties of the early childhood development measures used by World Bank to evaluate intervention impacts.

## Installation

```{r, eval = FALSE}
install.packages("remotes")
remotes::install_github("knickodem/WBdif")
library(WBdif)
```


## Features

Say you have a measure with both dichotomous and polytomous items and you want to know:

  1. Do any items show DIF by gender?
  2. What method should I use to detect DIF?
  3. What is the treatment effect of the intervention?
  4. Does the treatment effect change when accounting for biased (DIF-y) items?
  
**WBdif** has functionality to accommodate two approaches to answering these questions depending on your motivations and the nature of the measure (assessment/instrument/test).

### All-in-One {#aio}

The first approach requires only two functions (`dif_data_prep()` and `summary_report()`) and produces an HTML report summarizing the DIF analysis results from up to 4 methods along with tables and plots illustrating treatment effects with and without adjusting for item bias.

`dif_data_prep()` is a pre-processing step that organizes input information for use in other **WBdif** functions and conducts some checks on the suitability of the data for a DIF analysis. When investigating treatment effect robustness, `dif_data_prep()` is also where you specify:

 - whether the treatment effects are unconditional (i.e., only comparing treatment group to control group) or conditional on another variable (e.g., gender, race, age)
 - which standard deviation to use in the denominator of the treatment effect size estimation (`std.group`)
 - whether the data is clustered
 
`summary_report()` is then a wrapper around all of the other functions needed to run the analysis and produce the report (hence All-in-One). Nonetheless, the function offers a variety of options to customize the both the analysis and reporting. For more details on each option see the [Step-by-Step](#sbs) section. Given the All-in-One nature of `summary_report()`, the code may take a few minutes to run before producing the report.

**Conditional Treatment Effects**

Below we specify both `tx.group.id` and `dif.group.id` indicating that we are interested in estimating the treatment effects conditional on gender. The code below generates [this report](https://htmlpreview.github.io/?https://github.com/knickodem/WBdif/blob/master/DIF-Effects-Gender-MDAT-Language.html).

```{r, eval = FALSE}
data("mdatlang")

conditional <- dif_data_prep(item.data = mdatlang[5:ncol(mdatlang)],
                         tx.group.id = mdatlang$treated,
                         dif.group.id = mdatlang$gender,
                         cluster.id = mdatlang$clusterid,
                         std.group = NULL, # When NULL, the pooled standard deviation is used
                         na.to.0 = TRUE)

summary_report(dif.data = conditional,
             report.type = "dif.effects",
             report.title = "Gender DIF Effects on MDAT Language",
             measure.name = "MDAT Language",
             file.name = "DIF-Effects-Gender-MDAT-Language",
             dataset.name = "Malawi",
             methods = c("loess", "MH", "logistic", "IRT"),
             bias.method = "IRT",
             match.type  = "Total")
```

**Unconditional Treatment Effects**

When interested in unconditional effects, only one of `tx.group.id` or `dif.group.id` need to be supplied in `dif_data_prep()`.  `summary_report()` recognizes how the data are prepared and estimates the appropriate treatment effect. Thus, from an analysis standpoint, `dif.data` is the only argument that needs to be changed in `summary_report()`. Naturally, adjusting `report.title` helps clarify which analysis was conducted and and revising `file.name` prevents your previous reports from being overwritten. The code below generates [this report](https://htmlpreview.github.io/?https://github.com/knickodem/WBdif/blob/master/DIF-Effects-Tx-MDAT-Language.html).

```{r, eval = FALSE}
data("mdatlang")

unconditional <- dif_data_prep(item.data = mdatlang[5:ncol(mdatlang)],
                         tx.group.id = mdatlang$treated,
                         dif.group.id = NULL,
                         cluster.id = mdatlang$clusterid,
                         std.group = "Control", # "Control" is a value in mdatlang$treated
                         na.to.0 = TRUE)

summary_report(dif.data = unconditional,
             report.type = "dif.effects",
             report.title = "Tx DIF Effects on MDAT Language",
             measure.name = "MDAT Language",
             file.name = "DIF-Effects-Tx-MDAT-Language",
             dataset.name = "Malawi",
             methods = c("loess", "MH", "logistic", "IRT"),
             bias.method = "IRT",
             match.type  = "Total")
```

### Step-by-Step {#sbs}

If want to examine the components contributing to `summary_report()`, you can conduct the analysis and reporting one step at a time. The first recommended step is still `dif_data_prep()`, which operates in the same manner as the [All-in-One](#aio) approach. Once again we have specified both `tx.group.id` and `dif.group.id` to indicate we are interested in DIF by Gender and the conditional treatment effects.

```{r, message = FALSE, warning = FALSE}
library(WBdif)
data("mdatlang")
prepped <- dif_data_prep(item.data = mdatlang[5:ncol(mdatlang)],
                         tx.group.id = mdatlang$treated,
                         dif.group.id = mdatlang$gender,
                         cluster.id = mdatlang$clusterid,
                         std.group = NULL, # When NULL, the pooled standard deviation is used
                         na.to.0 = TRUE)
```

**DIF Analysis**

The next step is typically `dif_analysis()`, which evaluates DIF by the two groups in `dif.group.id` with up to 4 DIF methods:

 - LOESS regression - Generates plot of response curves for the two groups on each item. DIF is determined through visual inspection of differences in the groups' response curves.
 - Mantel-Haenszel (MH) Test - *only available for dichotomous items* - In a two-stage (Initial detection and Refinement stages) process, identifies biased items by conducting a Mantel-Haenszel test via `stats::mantelhaen.test()` for each item grouped by `dif.group.id` and stratified by raw total score (`match.type = "Total"`) or raw score removing the item under investigation (`match.type = "Rest"`).
 - Logistic regression - *only available for dichotomous items* - First conducts an omnibus test of DIF by comparing fit of no DIF, uniform DIF, and non-uniform DIF logistic regression models. If the model comparisons suggest DIF, biased items are identified by fitting a separate logistic regression model on each item during an initial detection stage and refinement stage.
 - Item Response Theory (IRT) - First conducts an omnibus test of DIF by comparing the fit of no DIF, uniform DIF, and non-uniform DIF 2PL IRT models via `mirt::multiGroup()`. If the model comparisons suggest DIF, biased items are identified by iteratively freeing parameters and testing model fit using `mirt::DIF()`.
 
Each method has its own function (`dif_loess()`, `dif_mh()`, `dif_logistic()`, and `dif_irt()`), which is called by `dif_analysis()`. One common issue in the MH method is low response rates for a certain group on a particular item at a given score. A common remedy is bin the scores, such as by deciles, as shown in the code below.
 
```{r, message = FALSE, warning = FALSE, results='hide'}
dif.analysis <- dif_analysis(dif.data = prepped,
                            methods =  c("loess", "MH", "logistic", "IRT"),
                            match.type = "Rest",
                            match.bins = seq(0, 1, by = .1))
```


If you only want a report of the DIF analysis and the treatment effect estimates are not of interest, a report including only the DIF results can be generated by setting `report.type = "dif.only"` in the `dif_report()` function. The report enables comparison of the DIF methods with one method highlighted via the `bias.method` argument ([example report](https://htmlpreview.github.io/?https://github.com/knickodem/WBdif/blob/master/DIF-Only-Gender-MDAT-Language.html)).

```{r, message = FALSE, warning = FALSE, eval=FALSE}
dif_report(dif.analysis = dif.analysis,
           report.type = "dif.only",
           report.title = "Gender DIF in MDAT Language",
           measure.name = "MDAT Language",
           file.name = "DIF-Only-Gender-MDAT-Language",
           dataset.name = "Malawi",
           bias.method = "IRT")
```


**Treatment Effect Estimation**

`effect_robustness()` estimates treatment effects using both a raw total score approach and an IRT model-based scoring approach. To do so, `dif_models()` is run first to gather and estimate the necessary IRT models. Typically, the inputs to `dif_models()` are the output from `dif_analysis()` and from which DIF method ("logistic", "MH", or "IRT") to extract the identified biased items. However, if you are interested in testing the impact of particular items, you can specify in `biased.items` a vector of the column indices in the data for the items you want to check along with `dif_data_prep()` object to `dif.data`.

The output from `dif_models()` is then used in `effect_robustness()` to estimate the treatment effects and their robustness in the presence of biased items. Which effects and how they are estimated is largely dictated by the information supplied to `dif_data_prep()`, which is passed to `effect_robustness()` from `dif_models()`. This includes:

 - unconditional or conditional treatment effects
 - the standard deviation used to standardize the effect size
 - whether the effect sizes and their standard errors are adjusted for the clustering of observations

```{r, message = FALSE, warning = FALSE, results='hide'}
    dif.models <- dif_models(dif.analysis = dif.analysis, dif.data = NULL, biased.items = "logistic")
    effect.robustness <- effect_robustness(dif.models = dif.models, std.group = prepped$std.group, irt.scoring = "WLE")
```

A report with just the robustness of treatment effect estimates to biased items, but no DIF analysis information, can be requested from `dif_report()` by setting `report.type = "effects.only"` and providing objects returned from `dif_models()` and `effect_robustness()`. More often, the DIF analysis results are also of interest, in which case `report.type = "dif.effects"` and the `dif_analysis()` object must be supplied. The format and data for the [report generated here](https://htmlpreview.github.io/?https://github.com/knickodem/WBdif/blob/master/Logistic-Gender-MDAT-Language.html) is the same as the Conditional Treatment Effects report in the [All-in-One](#aio) section, but the DIF methods used are different (logistic vs. IRT; rest scores vs total scores). Comparing the reports gives an indication of the impact of analysis decisions on conclusions regarding treatment effects.

```{r, message = FALSE, warning = FALSE, eval=FALSE}
dif_report(dif.analysis = dif.analysis,
           dif.models = dif.models,
           effect.robustness = effect.robustness,
           report.type = "dif.effects",
           report.title = "Gender DIF in MDAT Language",
           measure.name = "MDAT Language",
           file.name = "Logistic-Gender-MDAT-Language",
           dataset.name = "Malawi",
           bias.method = "logistic")
```


## Support and Suggestions

If you encounter any issues or have a suggestion for additional features, please file a [Github issue](https://github.com/knickodem/WBdif/issues) or contact us via email.