---
output:
  github_document:
    toc: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# DIFreport

**DIFreport is currently under construction. All information here is provisional and subject to change.**

## Summary

Development of **DIFreport** was motivated by questions regarding the psychometric properties of early childhood development assessments used as outcome measures in impact evaluations. This package provides methods for (a) evaluating differential item functioning (DIF), (b) checking whether DIF leads to biased standardized mean difference estimates on the unit-weighted total score or item response theory (IRT) based scoring methods, and (c) providing DIF-corrected estimates by either removing biased items or using an IRT models that adjust for DIF.

### Methods

The methods for evaluating DIF currently include semi-parametric regression (LOESS), the Mantel-Haenszel test, logistic regression, and multigroup IRT.  

When evaluating the impact of DIF, standardized mean differences can be estimated with respect to the primary grouping variable (e.g., gender, age, and socio-economic status) as well as with respect to a conditioning variable (e.g., intervention condition) The **DIFreport** documentation refers to these as unconditional and conditional standardized mean differneces, respectively.

### Workflows

**DIFreport** provides two general workflows, both of which are illustrated below. In the "[basic](#basic)" workflow, `summary_report()` is an omnibus function that automates the DIF analysis and, if there is any DIF detected, computes the standardized mean differences with and without adjusting for DIF. The results are summarized in an `rmarkdown` produced report. 

In the "[advanced](#advanced)" workflow, the steps automated by `summary_report()` can be implemented directly, which gives more control over which DIF analyses are conducted and which items are identified as biased.

The both workflows are currently under construction, but particularly the advanced workflow as the output returned by some functions are not very user-friendly. These outputs may be converted to S3 or S4 objects with methods (e.g., print, summary) to provide greater utility. 


**DIFreport** is in the alpha stage of development. Please report any bugs or suggestions by opening a [Github issue](https://github.com/knickodem/DIFreport/issues).

## Installation

```{r, eval = FALSE}
install.packages("remotes")
remotes::install_github("knickodem/DIFreport")
library(DIFreport)
```

# Workflows

## Dataset

Neuman, M.,  Ozler, B., & Fernald, L. (2013). *Protecting early childhood development in Malawi impact evaluation survey 2013, midline*. World Bank. https://doi.org/10.48529/94zr-ww41


## Basic {#basic}

The basic workflow requires only two functions: 

* `dif_prep()` is a pre-processing step that organizes input information for use in other **DIFreport** functions and conducts some checks on the data to preemptively address potential downstream issues. Additionally, different reports require different data preps. 

* `summary_report()` is a wrapper around all of the other functions needed to run the analysis and produce the report. The function offers a variety of options to customize the both the analysis and reporting. For more details on each option see the [Advanced workflow](#advanced) section. Given the all-in-one nature of `summary_report()`, the code will usually take several minutes to run before producing the report.

### Unconditional Standardized Mean Differences

If assessing DIF with respect to only the grouping variable of interest, and evaluating robustness of the unconditional standardized mean differences, only `dif.group` need be supplied to `dif_prep()`. If `cond.group` is supplied, then conditional standardized mean differences are estimated (see [Conditional Standardized Mean Differences](#conditional)). The function `summary_report()` recognizes how the data are prepared and estimates the appropriate standardized mean differences. 

The code below uses the built-in dataset "mdat" to generate [this report](https://htmlpreview.github.io/?https://github.com/knickodem/DIFreport/blob/master/README Example Reports/DIF-Effects-Tx-MDAT-Language.html). 

```{r, eval = FALSE}
# Load the example data (MDAT language assessment, collected in Malawi). 
data("mdat")

mdat_tx <- dif_prep(item.data = mdat[5:ncol(mdat)],
                               dif.groups = mdat$treated,
                               cluster = mdat$cluster,
                               ref.name = "Male", # "Male" is a value in mdat$gender
                               na.to.0 = TRUE)

summary_report(dif.data = mdat_tx,
               file.name = "DIF-Effects-Tx-MDAT-Language",
               report.type = "dif.effects",
               report.title = "MDAT Language: DIF by Treatment Condition",
               measure.name = "MDAT Language",
               dataset.name = "Malawi")
```

### Conditional Standardized Mean Differences {#conditional}

If assessing DIF with respect to a conditioning variable (e.g., treatment group) and evaluating robustness of the conditional standardized mean differences (e.g., for just girls, for just boys, and their difference) is of interest, then both `dif.groups` and `cond.groups`need to be supplied to `dif_prep()`.

The code below generates [this report](https://htmlpreview.github.io/?https://github.com/knickodem/DIFreport/blob/master/README Example Reports/DIF-Effects-Gender-MDAT-Language.html).

```{r, eval = FALSE}
mdat_gender <- dif_prep(item.data = mdat[5:ncol(mdat)],
                             dif.groups = mdat$gender,
                             cond.groups = mdat$treated,
                             cluster = mdat$cluster,
                             ref.name = "Male") 

summary_report(dif.data = mdat_gender,
               file.name = "DIF-Effects-Gender-MDAT-Language",
               report.type = "dif.effects",
               report.title = "MDAT Language: DIF by Gender",
               measure.name = "MDAT Language",
               dataset.name = "Malawi")

```


## Advanced {#advanced}

The advanced workflow shows what is going on under the hood of `summary_report()` and lets you have more control over which items are identified as biased in the report. In this section we work through the conditional standardized mean difference example in more detail. 

The first step is still `dif_prep()`, which operates in the same manner as the [Basic workflow](#basic). This step is repeated below. 


```{r, message = FALSE, warning = FALSE, eval=FALSE}
library(DIFreport)
data("mdat")

mdat_gender <- dif_prep(item.data = mdat[5:ncol(mdat)],
                             dif.groups = mdat$gender,
                             cond.groups = mdat$treated,
                             cluster = mdat$cluster,
                             ref.name = "Male") 
```

### DIF Analysis

The `dif_analysis()` function explores DIF with respect to the `dif.groups` using various methods. The goal is to arrive at an overall picture of the data that does not depend on the assumptions of any one particular method. Currently, 4 DIF methods are available for dichotomous data, but only semi-parametric regression and IRT can be used with multi-category ("polytomous") assessment items.  

 - LOESS regression - Generates plot of response curves for the two groups on each item. DIF is determined through visual inspection of differences in the groups' response curves.

 - Mantel-Haenszel (MH) Test - *currently only available for dichotomous items* -  Biased items are identified by conducting a Mantel-Haenszel test (`stats::mantelhaen.test()`) for each item by `dif.groups` table, stratified by either the unit-weighted total score (`match.type = "Total"`) or the unit-weighted total score computed after removing the item under investigation (`match.type = "Rest"`). Analysis proceeds in two stages, an initial "Purification" stage that tentatively identifies DIF items, and a second "Refinement" stage which removes the tentatively-identified items from the matching variable. The results of the second stage are considered more definitive. One common problem that arises with the MH test is when there are empty cells in the stratified two-way tables. A common remedy is to bin the stratifying variable (e.g., by deciles), as shown in the code below. 
 
 - Logistic regression - *only available for dichotomous items* - First conducts an omnibus test of DIF by comparing fit of no DIF, uniform DIF, and non-uniform DIF logistic regression models. If the model comparisons suggest DIF, biased items are identified by fitting two models for each item: one where all items parameters are held constant over groups, and one where the item in question interacts with `dif.groups`. Depending on the type of DIF identified in the omnibus test, the intercept (uniform DIF) or the intercept and slope (non-uniform DIF) may be interacted with `dif.groups`. The item-by-item testing uses the two-stage purification and refinement approach described in the MH Test.  
 
 - Item Response Theory (IRT) - 
 
Each method has its own low-level function (`dif_loess()`, `dif_mh()`, `dif_logistic()`, and `dif_irt()`), which are called by `dif_analysis()`. 
 
```{r, message = FALSE, warning = FALSE, results='hide', eval=FALSE}
dif.analysis <- dif_analysis(dif.data = mdat_gender,
                            dif.methods =  c("loess", "MH", "logistic", "IRT"),
                            match.type = "Total",
                            match.bins = seq(0, 1, by = .1)) # use deciles of matching variable in MH
```

```{r, eval=FALSE}
# View the biased items reported by "MH", "logistic", or "IRT" methods
options(knitr.kable.NA = '')
knitr::kable(biased_items_table(dif.analysis))
```

If the standardized mean difference estimates are not of interest, a report of the DIF analysis can be generated `dif_report()`. The report enables comparison of the DIF methods with one method highlighted via the `biased.items` argument ([example report](https://htmlpreview.github.io/?https://github.com/knickodem/DIFreport/blob/master/README Example Reports/DIF-Only-Gender-MDAT-Language.html)).

```{r, message = FALSE, warning = FALSE, eval=FALSE}
dif_report(dif.analysis = dif.analysis,
           file.name = "DIF-Only-Gender-MDAT-Language",
           report.title = "Gender DIF in MDAT Language",
           measure.name = "MDAT Language",
           dataset.name = "Malawi",
           biased.items = "IRT")
```


### DIF Impact

The last step in this workflow is to estimate standardized mean differences, with and without adjustments for DIF. standardized mean differences and their standard errors are computed using the method described by Hedges (2007), which is applicable to cluster-randomized designs as well as simple random samples. 

Standardized Mean Differences and their standard errors are reported for four different outcome variables.

* The unit-weighted total score computed with all items.

* The unit-weighted total score computed with DIF items omitted

* IRT scores computed using a model that constrains all items to have equal parameters over levels of `dif.groups`.

* IRT scores computed using a model that allows parameters of DIF-fy items to vary over levels of `dif.groups`.

The information supplied to `dif_prep()` (which is passed to `dif_impact()` through `dif.models`) provides additional details about how the standardized mean differences are computed 


* standardized mean differences are computed for `cond.groups`. If `cond.groups == dif.groups` unconditional standardized mean differences are reported. Otherwise standardized mean differences are computed condition on `dif.groups`. 
 
* The standard deviation used to standardize the effect size is given by `ref.name` if supplied.
 
* If `cluster` is provided, effect sizes and their standard errors are adjusted for a multi-stage sampling design in which `cluster` is the primary sampling unit. 
 
A report with just the robustness of standardized mean difference estimates to biased items, but no DIF analysis information, can be requested from `effect_report()`. More often, the DIF analysis results are also of interest, in which case use `dif_effect_report()` to include summaries of both the DIF analysis and the standardized mean difference robustness checks, such as the [report generated here](https://htmlpreview.github.io/?https://github.com/knickodem/DIFreport/blob/master/README Example Reports/Logistic-Gender-MDAT-Language.html)

```{r, message = FALSE, warning = FALSE, eval=FALSE}
dif_effect_report(dif.analysis = dif.analysis,
           dif.models = dif.models,
           effect.robustness = effect.robustness,
           file.name = "Logistic-Gender-MDAT-Language",
           report.title = "Gender DIF in MDAT Language",
           measure.name = "MDAT Language",
           dataset.name = "Malawi",
           biased.items = "logistic")
```


### References

Hedges, L. V. (2007). Effect Sizes in Cluster-Randomized Designs. Journal of Educational and Behavioral Statistics, 32, 341–370. https://doi.org/10.3102/1076998606298043.


