---
title: 'Summary of updates on "check" branch'
author: "PFH"
date: "Feb 5, 2020"
output:
  github_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Summary
New versions of the following files were created. A main theme of the changes was pulling stuff out of `effect_robustness` and writing stand alone functions for that stuff. 

  * newData-Prep.R: the function `dif_data_prep` now replaces `dif_prep`. It contains some stuff previously in `dif_analysis`: no.var.items, no.var.by.group.items, and poly.items.
  
  * newDIF-Analysis.R: Now includes `dif_models` which runs the `dif.mod` and, if necessary, the `no.dif.mod` previously in `effect_robustness`.
  
  * Hedges2007.R: replaces Est-Smd.R with an effect size function that has less input options.
  
  * new_Robusntess-Functions.R includes a stripped down version of `effect_robustness` that is basically
  just a wrapper for `hedges2007` as well as a bunch of helper functions that work on the output of `effect_robustness`. 
  
  * newPlot-Functions.R Minor updates to existing plots and added `effects_plot` previously in `effect_robustness`.
  
A second theme was to chain the functions so that, once the user specifices the data, the rest is on rails. If the user wants to run a conditional or an unconditional analysis, that is all specified in terms of the new `dif_data_prep` function. Once the data is prepped, that is the only input required by the user to produce a report. There are other ways to use the functions too but that was the target use case. 

A lot of variables were renamed along the way. I know this is a pain in the but I think it helps address the issues we talked about before (e.g., using "scale" in multiple different ways).

  * item.data replaces measure.data and data and similar variants
  * dif.group.id replaces dif.group
  * tx.group.id replaces tx.group
  * cluster.id replaces clusters
  * na.to.0 replaces na0
  * match.type replaces score.type
  * I am probably missing some 
  
 Note that I didn't do any renaming in functions that I did not edit, notably the low-level dif functions (e.g. loess). So changes of var.names is no universal yet. The probably don't need to be either. 

In summary, I did a lot of re-arranging in the "middle layer" of the project. What is left to do is plug the new output illustrated below into `dif_report`. I tried to make sure that the new outputs are comparably formatted to what was coming out of `effect_robusness`. The main difference is that the new outputs are formatted as list. Anyway, I hope it isn't too much of pain to patch up the report again...let me know if it is.  

# Example

Data prep is now via `dif_data_prep`. 

```{r, message=FALSE}
library(WBdif)
load("Bangladesh_Recoded.RData")
midline <- bang.recode[bang.recode$line == "Mid",]

i <- 1
midline <- bang.recode[bang.recode$line == "Mid",]
i <- 1

# data prep is now via dif_data_prep
dif.data <- dif_data_prep(item.data = midline[domain.items[[i]]],
                         tx.group.id = midline$tx,
                         dif.group.id = midline$gender,
                         cluster.id = NULL,
                         na.to.0 = F)
```

`dif_analysis` now takes `dif.data` as input rather than components thereof. 

```{r, results='hide'}
dif.analysis <- dif_analysis(dif.data = dif.data,
                             methods = c("loess", "IRT"))
```
```{r}
names(dif.analysis)
```
If no biased items are detected by `dif.analysis`, we should skip every thing that follows and just produce the report that says there was no DIF (I realize this is a change from before). 

If there are biased items, `dif_models` runs the `dif.mod` identified by `dif.analysis`. This previously happened in `effect_robustness`. It also runs the `no.dif.mod` if needed, or just plucks it out of `dif.analysis`. 

The argument `biased.items` can be one of `c("IRT, "MH", "logistic")` or a numeric vector indicating which columns from `item.data` to treat as biased. 

All subsequent functions require `dif.analysis`, so, if it needs to be forced to run even when there was no `dif.analysis` reported no DIF (e.g., for debugging), manually enter a numeric vector for `biased.items`. 

```{r, results='hide'}
dif.models <- dif_models(dif.analysis, biased.items = c(1,2,3))
```
```{r}
names(dif.models)
```

I spend a lot of time arriving back at the same plots we had in the first place...sigh

```{r}
bias.plots <- bias_plots(dif.models)
gridExtra::grid.arrange(bias.plots[[1]], bias.plots[[2]])
```

The funciton `effect_robustness` is now just a wrapper for `hedges2007`, which replaces `est_smd` and `smd_wrapper`. The output is a `list` whose components are each like the `effects.data` object that was internal to `effects_robustness`. Functions for computing effect plots, tables, and coeff alpha are now external to `effect_robustness`. But, it still takes a while to run due to calling `mirt::fscores` ... 

```{r}
effects.list <- effect_robustness(dif.models)
names(effects.list)
```

The effect size plots previously in `effect_robustness` are now called via `effects_plot` in the `newPlot-Functions.R` file

```{r}
(effects.plots <- lapply(effects.list, effects_plot))
```

The effect tables previously in `effect_robustness` is called via `effects_table`. The format is somewhat different, as I dropped the reliability adjusted effect sizes for the total scores and included the standard errors from `hedges2007`. 

```{r}
effects.tables <- lapply(effects.list, effects_table)
lapply(effects.tables, knitr::kable)
```

Coefficient alpha is also now external to `effect_robustness`. This is just a wrapped on `est_alpha`.

```{r}
alphas.list <- coeff_alpha(dif.models)
lapply(alphas.list, knitr::kable)
```

The function `effects_table` has an optional argument to include the alphas. 

```{r}
effects.tables.with.alpha <- mapply(effects_table, effects.list, alphas.list)
lapply(effects.tables.with.alpha, knitr::kable)
```

As I mentioned above, I didnt try to sort out how to put this into the report. I am hoping the individual pieces shown above will be easy to work with but let me know if not. I will work on updating documentation over the next few days.  
