---
title: "WB Project Decisions"
output:
  html_document:
    toc: yes
always_allow_html: yes
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Identifying data errors

**EGMA_number_recognition.Endline**: The stage1 MH test produced a statistic, df, and p-value, but no estimate or CI.

```{r data error check}
library(dplyr)
check <- MalawiData[384:403] %>%
  tidyr::gather(Item, Response) %>%
  group_by(Item, Response) %>% summarize(n = n()) %>%
  mutate(r = n())
```

 - recog4_3 has a value of 3 in one cell (instead of all 0s and 1s)
 - recog12_3 has a value of 2 in one cell
 - recog15_3 has a value of 9 in one cell
 

# Alternative approaches

Comparing results from our code to using the difR package as a sensitivity check


# Decisions, assumptions, and defaults

I am currently organizing this by function, which should be helpful if this transformed to an r package

## WB_Data_Prep

 - Assumes the last 2 characters in the item (column) names can be removed
 - Only removes cases with missing data on all items or in the grouping variable
 - Recodes NAs to 0
   - Thus, scores could be a product of motivation rather than ability; in other words, the analysis equates low motivation to low ability  



## DIF_analysis

 - Assumes the grouping variable (groupvec) is a factor with 2 levels


