
```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(options(knitr.kable.NA = ''))

```

---
title: "`r report.title`"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
---

# Summary

```{r dataset, results='asis'}
if(dataset.name != "dataset"){
  cat("**Dataset:**", dataset.name)
  cat("\n\n")
}
```
  

**Measure:** `r if(measure.name == "measure") outputs$item.name.range else paste0(measure.name,  " (", outputs$item.name.range, ")")`

**Comparison groups:** `r if(length(outputs$dif.groups) == 2) paste(outputs$dif.groups, collapse = " and ") else paste(outputs$dif.groups, collapse = ", ")`

**Items with no variance:** `r outputs$no.var.items`

**Items with no variance within comparison groups:** `r outputs$no.var.by.group.items`

**DIF detection method:** `r if(biased.items == "IRT") biased.items else paste(biased.items, "using", match.type, "Scores")`

**Total number of biased items:** `r paste(n.biased, "out of", outputs$n.items)`

**Type of DIF:** `r dif.type`

```{r direction, results='asis', ft.align="left"}
if(dif.type == "uniform"){
  if(n.biased > 0){
  
  cat("*Number of items uniformly biased toward each group*")
  cat("\n\n")
  toward.flex <- format_flex(toward)
  cat(knitr::knit_print(toward.flex))
  cat("\n\n")
  }
}
```
  

```{r No Var, results='asis'}
if(outputs$no.var.items != "none"){
  
  cat("\n\n")
  cat("^a^ Items were not included in any DIF analysis and were not used to compute 
      treatment effects.")
  cat("\n\n")
}


if(outputs$no.var.by.group.items != "none"){
  
  cat("\n\n")
  cat("^b^ Items were not included in the IRT-based DIF analysis and were not used 
      to compute IRT-based treatment effects.")
  cat("\n\n")
}

```


# DIF Detection Methods

*Note.* Logical values in bias columns are based on p-values adjusted with the Benjamini & Hochberg (1995) method rather than directly on the p-value columns.

## Biased Item Summary

```{r bi summary, results = 'asis', ft.align="left"}

if(nrow(biased.items.table) != 0){
    bi.flex <- format_flex(biased.items.table)
    cat(knitr::knit_print(bi.flex))
    cat("\n\n")
    cat("&nbsp;  ")
    cat("*Note:* X = method determined item to be biased")
    cat("\n\n")
} else {
    cat("Item bias was not detected by any of the selected methods")
    cat("\n\n")
}

```



```{r loess, results='asis', warning=FALSE, fig.width = 10, fig.height = ceiling(outputs$n.items / 6)*2 + 1}

if(!is.null(dif.analysis$loess)){
  cat("## Graphical Analysis via loess")
  cat("\n\n")
  
  print(dif.analysis$loess$plot)
  
}
```



```{r MH, results='asis', ft.align="left"}

if(!is.null(dif.analysis$MH)){
  
  cat("## Mantel-Haenszel")
  cat("\n\n")
    
  if(class(dif.analysis$MH$item.level) == "character"){
        
    cat(dif.analysis$MH$item.level)
    cat("\n\n")
    
  } else {
    
    MHtab <- format_flex(dif.analysis$MH$item.level, bold.bias = "item")
    cat(knitr::knit_print(MHtab))
    cat("\n\n")
  }
}
```




```{r Logistic Global, results='asis', ft.align="left"}
if(!is.null(dif.analysis$logistic)){
  
  cat("## Logistic Regression")
  cat("\n\n")
  
  cat("#### Model Comparisons")
  cat("\n\n")

  logglobtab <- format_flex(dif.analysis$logistic$global.level, bold.bias = "global")
  cat(knitr::knit_print(logglobtab))
  cat("\n\n")
  
  cat("&nbsp;  ")
  cat("\n\n")
  cat("#### Item-level Tests")
  cat("\n\n")
  
  if(class(dif.analysis$logistic$item.level) == "character"){
    
    cat(dif.analysis$logistic$item.level)
    cat("\n\n")
        
  } else {
    
    logitemtab <- format_flex(dif.analysis$logistic$item.level, bold.bias = "item")
    cat(knitr::knit_print(logitemtab))
    cat("\n\n")
  }

}
```


```{r IRT Global, results='asis', ft.align="left"}

if(!is.null(dif.analysis$IRT)){

  cat("## Item Response Theory")
  cat("\n\n")
  
  cat("#### Model Comparisons")
  cat("\n\n")
  
  irtglobtab <- format_flex(dif.analysis$IRT$global.level, bold.bias = "global")
  cat(knitr::knit_print(irtglobtab))
  cat("\n\n")
 
  cat("&nbsp;  ")
  cat("\n\n")
  cat("#### Item-level Tests")
  cat("\n\n")
  
  if(class(dif.analysis$IRT$item.level) == "character"){
    
    cat(dif.analysis$IRT$item.level)
    cat("\n\n")
        
  } else {
    
    irtitemtab <- format_flex(dif.analysis$IRT$item.level, bold.bias = "item")
    cat(knitr::knit_print(irtitemtab))
    cat("\n\n")
  }
}

```