
```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(options(knitr.kable.NA = ''))

```

---
title: "`r report.title`"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
---

# Summary

```{r dataset, results='asis'}
if(dataset.name != "dataset"){
  cat("**Dataset:**", dataset.name)
  cat("\n\n")
}
```
  

**Measure:** `r if(measure.name == "measure") outputs$item.name.range else paste0(measure.name,  " (", outputs$item.name.range, ")")`

**Comparison groups:** `r if(length(outputs$dif.groups) == 2) paste(outputs$dif.groups, collapse = " and ") else paste(outputs$dif.groups, collapse = ", ")`

**Items with no variance:** `r outputs$no.var.items`

**Items with no variance within comparison groups:** `r outputs$no.var.by.group.items`

**Biased items:** `r paste(names(inputs$item.data)[dif.models$biased.items], collapse = ", ")`


```{r No Var, results='asis'}
if(outputs$no.var.items != "none"){
  
  cat("\n\n")
  cat("^a^ Items were not included in any DIF analysis and were not used to compute 
      treatment effects.")
  cat("\n\n")
}


if(outputs$no.var.by.group.items != "none"){
  
  cat("\n\n")
  cat("^b^ Items were not included in the IRT-based DIF analysis and were not used 
      to compute IRT-based treatment effects.")
  cat("\n\n")
}

```


# Treatment Effect Robustness

```{r bias.plots, results='asis', message=FALSE, fig.width = 5, fig.height = 6}

gridExtra::grid.arrange(bias.plots[[1]], bias.plots[[2]])

```

"Bias Omitted" treatment effect estimates adjusted for bias in `r paste(names(inputs$item.data)[dif.models$biased.items], collapse = ", ")`.

```{r Treatment_Effects, results='asis', warning=FALSE, ft.align="left", fig.width = 5, fig.height = 4}

for(i in 1:length(effects.tables)){
  cat("\n\n")
  cat("##", names(effects.tables)[[i]])
  cat("\n\n")
  et.flex <- format_flex(effects.tables[[i]])
  cat(knitr::knit_print(et.flex))
  cat("\n\n")
  cat("&nbsp;  ")
  
  print(effects.plots[[i]])
  
}
```

